---
title: "Homework for Network Analysis Workshop"
author: "Min Hee Seo"
date: "August 21st, 2018"
output: pdf_document
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE)
```

\section{Exercise 1. Nigeria Data}

```{r ex1, eval=T, echo=T, warning=F, message=F, tidy=TRUE, tidy.opts=list(width.cutoff=60)}
# create network matrix by year and all years

# setup the working directory
setwd("/Users/minheeseo/Dropbox/Classes/2018_Classes/Network/network2018_hw1/")
# clean the workspace
rm(list=ls())
# loading data and R packages
library(igraph)
library(network)
load("nigeria.rda")
nigeria$sender<- gsub("\n", " ", nigeria$sender)
nigeria$receiver <- gsub("\n", " ", nigeria$receiver)

# create list where the length of list is time span
network.mat <- vector("list",length(unique(nigeria$year))+1)
names(network.mat) <- unique(nigeria$year)

time <- unique(nigeria$year)
for(t in 1:length(time)){
  slice <- NULL
  empty.mat <- NULL
  country.senter <- country.receiver <- c()
  slice <- nigeria[nigeria$year==time[t],]
  country.sender <- unique(slice$sender)
  empty.mat <-matrix(0, length(country.sender), length(unique(slice$receiver)))
  empty.mat <- as.data.frame(empty.mat)
  rownames(empty.mat) <- country.sender
  colnames(empty.mat) <- unique(slice$receiver)
  for(i in 1:length(country.sender)){
    country.receiver <- unique(slice$receiver[slice$sender==country.sender[i]])
    for(j in 1:length(country.receiver)){
      empty.mat[rownames(empty.mat)==country.sender[i], colnames(empty.mat)==country.receiver[j]] <- slice$conflict[slice$sender==country.sender[i] & slice$receiver==country.receiver[j]]
    }
  }
    network.mat[[t]] <-empty.mat
}
# network.mat list contains 17 matrix each one for each year
country.sender <- unique(nigeria$sender)
empty.mat <-matrix(0, length(country.sender), length(unique(nigeria$receiver)))
empty.mat <- as.data.frame(empty.mat)
rownames(empty.mat) <- country.sender
colnames(empty.mat) <- unique(nigeria$receiver)
for(i in 1:length(country.sender)){
  country.receiver <- unique(nigeria$receiver[nigeria$sender==country.sender[i]])
  for(j in 1:length(country.receiver)){
    empty.mat[rownames(empty.mat)==country.sender[i],colnames(empty.mat)==country.receiver[j]] <- sum(nigeria$conflict[nigeria$sender==country.sender[i] & nigeria$receiver==country.receiver[j]])
  }
}
names(network.mat)[18] <- "All Conflict"
network.mat[[18]] <- empty.mat
```


```{r ex1plot, eval=T, echo=T, warning=F, message=F}
# plot network by each year and all years

myblue <- rgb(red = 0, green = 0, blue = 1, alpha = .5)
mypink <- rgb(red = 1, green = 0, blue = 0, alpha = .25)

par(mfrow=c(2, 2), mar=c(0,0.2,1,0.2))
for(i in 1:4){
  g <- NULL
  g = graph_from_adjacency_matrix(as.matrix(network.mat[[i]]),
                                mode='directed', weighted=TRUE, diag=F)
  tiesSum = apply(g[], 1, sum)
# condition size based on # of ties
  V(g)$size <- (tiesSum+0.5)*6
# only label if # ties greater than 10
  V(g)$label <- ifelse( tiesSum>1, V(g)$name, NA )
  V(g)$label.cex <- 0.6
  plot(g,main=paste("Year:", names(network.mat)[i]),
     vertex.label=V(g)$label, 
     vertex.size=V(g)$size,
     edge.width=E(g)$weight,
     vertex.color =mypink, # change color of nodes
     vertex.label.color = "black", # change color of labels
     edge.curved=.25, # add a 25% curve to the edges
     edge.color=myblue, # change edge color to grey  
     layout=layout_with_fr)
}

```


```{r ex1plot2, eval=T, echo=F, warning=F, message=F}

par(mfrow=c(2, 2), mar=c(0,0.2,1,0.2))
for(i in 5:8){
  g <- NULL
  g = graph_from_adjacency_matrix(as.matrix(network.mat[[i]]),
                                mode='directed', weighted=TRUE,diag=F)
  tiesSum = apply(g[], 1, sum)
# condition size based on # of ties
  V(g)$size <- (tiesSum+0.5)*6
# only label if # ties greater than 10
  V(g)$label <- ifelse( tiesSum>1, V(g)$name, NA )
  V(g)$label.cex <- 0.6
  plot(g,main=paste("Year:", names(network.mat)[i]),
     vertex.label=V(g)$label, 
     vertex.size=V(g)$size,
     edge.width=E(g)$weight,
     vertex.color =mypink, # change color of nodes
     vertex.label.color = "black", # change color of labels
     edge.curved=.25, # add a 25% curve to the edges
     edge.color=myblue, # change edge color to grey  
     layout=layout_with_fr)
}
```

```{r ex1plot3, eval=T, echo=F, warning=F, message=F}
par(mfrow=c(2, 2), mar=c(0,0.2,1,0.2))
for(i in 9:12){
  g <- NULL
  g = graph_from_adjacency_matrix(as.matrix(network.mat[[i]]),
                                mode='directed', weighted=TRUE,diag=F)
  tiesSum = apply(g[], 1, sum)
# condition size based on # of ties
  V(g)$size <- (tiesSum+0.5)*6
# only label if # ties greater than 10
  V(g)$label <- ifelse( tiesSum>1, V(g)$name, NA )
  V(g)$label.cex <- 0.6
  plot(g,main=paste("Year:", names(network.mat)[i]),
     vertex.label=V(g)$label, 
     vertex.size=V(g)$size,
     vertex.color=mypink,
     edge.width=E(g)$weight,
     vertex.label.color="black", 
     edge.curved=.25, 
     edge.color=myblue,
     layout=layout_with_fr)
}
```

```{r ex1plot4, eval=T, echo=F, warning=F, message=F}
par(mfrow=c(2, 2), mar=c(0,0.2,1,0.2))
for(i in 13:16){
  g <- NULL
  g = graph_from_adjacency_matrix(as.matrix(network.mat[[i]]),
                                mode='directed', weighted=TRUE,diag=F)
  tiesSum = apply(g[], 1, sum)
# condition size based on # of ties
  V(g)$size <- (tiesSum+0.5)*6
# only label if # ties greater than 10
  V(g)$label <- ifelse( tiesSum>1, V(g)$name, NA )
  V(g)$label.cex <- 0.6
  plot(g,main=paste("Year:", names(network.mat)[i]),
     vertex.label=V(g)$label, 
     vertex.size=V(g)$size,
     edge.width=E(g)$weight,
     vertex.color=mypink, # change color of nodes
     vertex.label.color="black", # change color of labels
     edge.curved=.25, # add a 25% curve to the edges
     edge.color=myblue, # change edge color to grey  
     layout=layout_with_fr)
}

```


```{r ex1plot5, eval=T, echo=F, warning=F, message=F}
par(mfrow=c(2, 2), mar=c(0,0.2,1,0.2))
g <- NULL
g = graph_from_adjacency_matrix(as.matrix(network.mat[[17]]),
                                mode='directed', weighted=TRUE,diag=F)
tiesSum = apply(g[], 1, sum)
# condition size based on # of ties
V(g)$size <- (tiesSum+0.5)*6
# only label if # ties greater than 10
V(g)$label <- ifelse( tiesSum>1, V(g)$name, NA )
V(g)$label.cex <- 0.6
plot(g,main=paste("Year:", names(network.mat)[17]),
     vertex.label=V(g)$label, 
     vertex.size=V(g)$size,
     edge.width=E(g)$weight,
     vertex.color =mypink, # change color of nodes
     vertex.label.color = "black", # change color of labels
     edge.curved=.25, # add a 25% curve to the edges
     edge.color=myblue, # change edge color to grey  
     layout=layout_with_fr)
g <- NULL
g = graph_from_adjacency_matrix(as.matrix(network.mat[[18]]),
                                mode='directed', weighted=TRUE,diag=F)
tiesSum = apply(g[], 1, sum)
# condition size based on # of ties
V(g)$size <- tiesSum*1/2.5
# only label if # ties greater than 10
V(g)$label <- ifelse( tiesSum>1, V(g)$name, NA )
V(g)$label.cex <- 0.45
plot(g,main=names(network.mat)[18],
     vertex.label=V(g)$label, 
     vertex.size=V(g)$size,
     edge.width=E(g)$weight,
     vertex.color=mypink, # change color of nodes
     vertex.label.color = "black", # change color of labels
     edge.curved=.25, # add a 25% curve to the edges
     edge.color=myblue, # change edge color to grey  
     layout=layout_with_fr)


```
\section{Exercise 2. Measurements \& Community detection}
\subsection{a}
We can measure influence with the degree of nodes. For example, we can count a total number of edges connected to a node to estimate the influence of an actor. By comparing network plot (above) and the degree (below), I am quite certain that I found an influential actor using degree. Additionally, I computed eigenvector centrality to find a node that is linked to other important nodes. When I compare actors who have the most degree and eigenvector centrality, they were almost the same. I report the most influential actor for each year and all years (in terms of the number of degree).

```{r degree, eval=T, echo=T, warning=F, message=F}
# to find an influential actor for each year

# create an empty list
influence.actor <- vector("list",18)
names(influence.actor) <- unique(nigeria$year)
eigen.actor <- vector("list",18)
names(eigen.actor) <- unique(nigeria$year)
# loop to find the maximum degree and eigenvector centrality
for(t in 1:17){
  temp <- graph.adjacency(as.matrix(network.mat[[t]]), mode='directed', weighted=TRUE,diag=F)
  degree <- igraph::degree(temp)
  eigen <- eigen_centrality(temp, directed=TRUE)$vector
  names(eigen) <- as.character(gsub("\n", " ", names(eigen)))
  names(degree) <- as.character(gsub("\n", " ", names(degree)))
  influence.actor[[t]] <- degree[which(degree==max(degree))]
  eigen.actor[[t]] <- eigen[which(eigen==max(eigen))]
}
# to find an influential actor overall
names(influence.actor)[18] <- "All Conflict"
names(eigen.actor)[18] <- "All Conflict"
temp <- graph.adjacency(as.matrix(network.mat[[18]]),mode='directed', weighted=TRUE,diag=F)
degree <- igraph::degree(temp)
eigen <- eigen_centrality(temp, directed=TRUE)$vector
names(eigen) <- as.character(gsub("\n", " ", names(eigen)))
names(degree) <- as.character(gsub("\n", " ", names(degree)))
influence.actor[[18]] <- degree[which(degree==max(degree))]
eigen.actor[[18]] <- eigen[which(eigen==max(eigen))]

influence.actor # for each year and all years
eigen.actor[[18]] # for all years

```



\subsection{b}
```{r blockmodel, eval=T, echo=T, warning=F, message=F}
library(sna)

# gauge data
g <- network(network.mat[[18]], directed=T)
eclusts <- equiv.clust(g)
plot(eclusts, hang=-1) # seems like there are 8 groups

# running the block model using cross-validation

# save the node classification from each run

# out of sample cross validation to validate "k"

# report ROC statistics from each model

```


\section{Exercise 3. ERGMs}
```{r ergms, eval=T, echo=T, warning=F, message=F}
library(statnet)
# for conflicts in 2004
set.seed(688346)
nigeria <- as.matrix(network.mat[[5]])
nigeria <- as.network.matrix(nigeria)
ergm.network <- ergm(nigeria ~ edges + mutual + idegree1.5)
mcmc.diagnostics(ergm.network)
```